### 2013-09-03 report/prefAttachBrief.tex
CPI successfully running, with a number of added output files to visualize the results. started work on a brief detailing current progress.

\documentclass[11pt]{article}
\setlength{\textheight}{240mm}
%\setlength{\voffset}{0mm}
\addtolength{\topmargin}{-30mm}
\setlength{\textwidth}{155mm}
\setlength{\oddsidemargin}{5mm}
\usepackage{graphicx, subfig}
\pagestyle{plain}
\begin{document}
\title{Preferential Attachment Coarse Projective Integration}
\author{Alexander Holiday\vspace{-2ex}}
\date{09/01/2013}
\maketitle
\section*{Introduction}
The preferential attachment model, detailed in \cite{balasz:rsa12}, was implemented in C++. The system naturally converges to a limiting ``graphon'' (see \cite{lovasz:jcombth06}), and, in order to increase the rate of convergence, coarse projective integration was implemented. This brief outlines the model dynamics and reviews current simulation progress.
\section*{Preferential Attachment Model}
This vertex- and edge-conservative model begins with an Erd\H{o}s-R\'{e}nyi random graph in which $m$ edges are randomly assigned to $n$ vertices.