### 2015-09-08 report/pa_diff.tex
first attempt with magit, don't fully understand what's happening. adding rho/kappa embedding file to the mix, switched up organization of files.
merge conflict arose from upper- vs. lower-case legend labels. silly.

\documentclass[11pt]{article}
\usepackage{graphicx, subcaption}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
\graphicspath{ {./figs/} }
\pagestyle{plain}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist 
\setlength\parindent{0cm}
\begin{document}
\title{\vspace{-5mm}Differences between canonical preferential attachment model and Bal\'{a}zs'}
\maketitle

The canonical preferential attachment model, described in the attached paper and called the ``Bar\'{a}basi-Albert model'', describes a method of \textbf{constructing} a network that will have a power-law degree distribution. The method is simple:

\begin{enumerate}
\item Begin with a complete graph on $m_0$ vertices
\item Add an additional $m$ vertices which will be connected to $k \leq m_0$ of the existing vertices with probability $P_i = \frac{d_i}{\sum\limits_j d_j}$ where $d_i$ is the degree of $v_i$
\end{enumerate}

Bal\'{a}zs' model, fully called the ``edge-conservative preferential attachment model'' describes, of course, a method of \textbf{evolving} a network. At each step, the probability that an edge is added to vertex $i$ is given by $P_i = \frac{d_i + \kappa}{\sum\limits_i d_i + n \kappa}$. \\ \\ \\

{\huge Thus when they hear ``preferential attachment'', people may think of a network \textbf{construction} algorithm, and may be familiar with $P_i = \frac{d_i}{\sum\limits_j d_j}$ instead of $P_i = \frac{d_i + \kappa}{\sum\limits_j d_j + n \kappa}$.}

%% \begin{changemargin}{0.0cm}{0.0cm}
%% \centering

%% \include{balazs_briefing_figs}

%% \end{changemargin}

\end{document}