### 2015-09-08 report/paper_outline.tex

\documentclass[12pt]{article}
\usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
\usepackage[top=1in, bottom=1in, left=0.5in, right=0.5in]{geometry}
\usepackage[center]{titlesec}
\graphicspath{ {./figs/paper/} }
\pagestyle{plain}
\renewcommand\thesection{\Roman{section}}
\renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
\setlength{\parskip}{0ex}
\def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
\let\endchangemargin=\endlist

\title{\vspace{-2cm}Equation-free analysis of a dynamically evolving multigraph}
\author{Alexander Holiday,$^1$ Bal\'{a}zs R\'{a}th,$^2$ Ioannis G. Kevrekidis$^{1,3}$ \\ \\
  $^1$\textit{Department of Chemical and Biological Engineering}, \\
  \textit{Princeton University, Princeton, New Jersey 08544, USA} \\ \\
  $^2$\textit{Department of Mathematics}, \\
  \textit{The University of British Columbia, Vancouver, British Columbia V6T 1Z4}
}
\date{}

\begin{document}
\maketitle
\begin{singlespace}
\vspace{-5mm}
  \begin{enumerate}
  \item Introduction
    \begin{enumerate}
    \item Examples of multigraphs
    \item Difficulty of analysis
    \item Benefits of equation-free methods
    \end{enumerate}
  \item Model Description
    \begin{enumerate}
    \item Step-by-step dynamics
    \item Timescales of degree distribution evolution
    \end{enumerate}
  \item EF Modeling
    \begin{enumerate}
    \item Coarse-graining via degree distribution
    \end{enumerate}
    \begin{enumerate}
    \item Restriction
      \begin{enumerate}
      \item Multigraph $\rightarrow$ sorted degree distribution (direct evaluation)
      \item Sorted degree distribution $\rightarrow$ fitted polynomial coefficients (constrained least-squares)
      \end{enumerate}
    \end{enumerate}
    \begin{enumerate}
    \item Lifting
      \begin{enumerate}
      \item Fitted polynomial coefficients $\rightarrow$ sorted degree distribution  (evaluation of polynomial)
      \item Sorted degree distribution $\rightarrow$  multigraph (modified Havel-Hakimi)
      \end{enumerate}
    \item Coarse-projective integration
    \end{enumerate}
  \item Newton-GMRES
  \item \textbf{Algorithmic coarse-graining}
    \begin{enumerate}
      \item \textbf{PCA}
        \begin{enumerate}
          \item \textbf{Algorithm overview}
          \item \textbf{Application to degree distribution}
        \end{enumerate}
      \item \textbf{DMAPS}
        \begin{enumerate}
          \item Algorithm overview
          \item Graph distances
          \item \textbf{Application to different timescales}
        \end{enumerate}
    \end{enumerate}
  \end{enumerate}

Unresolved questions for Bal\'{a}zs: Is there any property of the evolving multigraph that, in general, changes on the shorter $n^2$ timescale?

\end{singlespace}

%% \bibliographystyle{abbrv}
%% \bibliography{$HOME/Documents/bibTex/library}

\end{document}